Luminance constancy under fixed geometry

### 2018-05-11 Draft4/draft4.tex
DHB finished with this pass.  Onto JB.

Figure~\ref{fig:barGraphs} summarizes our main findings.
It plots the overall relative root mean squared estimation error (relative RMSE) for each condition. 
In Condition 1, where only the relative target object spectrum varies, luminance constancy poses an easy computationsl problem,
and performance is essentially perfect for both the baseline linear regression and AMA methods.
In Condition 2, where the spectral power distribution of the illumination also varies, the problem of luminance constancy is more difficult.
Indeed, neither of our methods performs well when they directly act on the cone excitations.
Performance with both methods improves greatly when the cone excitations are contrast normalized. 
Indeed, AMA performance for this case is, as in Condition 1, essentially perfect.
Linear regression does less well, in part because it only has access to cone excitations at image locations
corresponding to the target object.
Condition 2, however, does not include scene-to-scene variation in the background surface reflectances,
as occurs in natural viewing.
The results for Condition 3 show performance when we add such variation.
Performance here represents the overall level of luminance constancy that our methods achieve for the
most realistic condition that we tested.
In this case, AMA-based estimates of target object LRV are, on average, within 13\% (relative RMSE) of the true value.
Note that introduction of variation in the background surfaces makes the luminance constancy problem substantially harder,
because variation in the background surfaces reduces the reliability of the light reflected from these
surfaces as a cue to the spectral power distribution of the illuminant.

@@

As we noted in the introduction, supervised learning requires large labeled datasets and these are not readily available 
for the study of how to estimate correlates of object surface reflectance.
In particular, to obtain ground truth information about the reflectance of objects in a natural image, it is necessary to
make an independent of the illumination impinging on each object.
This can be done by inserting a small number of discretely spaced reflectance standards into the scene, but 
interpolating to locations between such standards requires strong assumptions.
The use of rendered images allows us to work with large numbers of labeled images where the reflectance
and impinging illumination are known precisely at every pixel.

A second advantage of using rendered images is that we can control the variation in distinct
scene factors that might affect the difficulty of the estimation problem.
This allows study of the individual effect of each factor as well as their joint effect.
Here, we exploited this advantage to systematically introduce variation in the relative target object reflectance
spectrum, the spectrum of the illumination, and the reflectance spectrum of the background objects.
This allowed us to quantify how each of these variations limits LRV estimation.
We believe this basic approach can be extended to include parametric control over the amount
of variation of different factors as well as to include variation in additional factors (see below).

At the same time, it is important to recognize that limitations of the approach.
Virtual rendered scenes are not guaranteed to capture all of the task relevant
variation that occurs in real scenes.
Indeed, even casual inspection of our images reveals that they are rendered and
not real.
Computer graphics is getting better, however, and we expect that the gap between
virtual and natural image databases will steadily close.
Indeed, carefully constructed graphics images are now quite difficult to differentiate
from real images.

Our use of virtual images relies on datasets that characterize the statistical variation
in object reflectance and natural illumination spectra.
We believe these datasets provide reasonable approximations.
None-the-less, they can be extended and improved.
For example, the dataset of illuminations does not include the artificial illuminations
that we now live much of our lives under.
The surface reflectance datasets captures constraints on naturally occurring reflectance
functions, but did not sample the environment in a manner that was intended to 
characterize the relative frequency of occurrence of different reflectance spectra.
Future work with virtual image datasets will benefit from targeted efforts to characterize
the natural statistics of the scene components that are varied.
Under the assumption that combination of different scene factors (e.g. surface reflectances
and illuminations) is independent, the virtual image approach modularizes the study
of variation in natural scenes.
%Supervised learning requires large labeled data sets.  Such data sets are not readily available for the study of color constancy. Although there are databases of calibrated color images, these do not provide ground truth information about surface reflectance and illuminantion at each image location \cite{ChakrabartiHyperspectral,NascimentoFoster2016,ParragaHyperspectralData,TkacikUpennHypersepctralData,skauli2013collection,olmos2004biologically}. There are some smaller databases consisting of images of posed scenes where object surface reflectances are measured individually (\citeNP{funt1988color,ciurea2003large}; \citeauthor{davidPennHyperspectral}), but these are not large enough to drive supervised learning.

\subsection{Future Directions}
In the work presented here, we incorporated spectral variation into our statistical model
of naturalistic scenes.
It is natural to begin with spectral variation, because this variation is at the heart of
what makes luminance constancy a rich computational problem.
In natural scenes, however, there are other sources of variation that add additional
richness.
These include variation in non-spectral properties of objects and lighting in the scene, including
object texture, material, and shape as well as lighting geometry.
The methods we developed here may be generalized to study the effects of variation
in these factors.
The basic approach would be to incorporate the variation into the generation of the
scenes and again learn estimators from the corresponding labeled images.
A challenge for this approach will be to thoughtfully control the increase in problem
complexity both to keep compute time feasible and to ensure that it is possible to
extract meaningful insight from the results.

Luminance constancy is a special case of the more general problem of color constancy.
The approach we developed here could easily be generalized to the estimation of additional
surface reflectance descriptors, such as object hue or chroma.
As with LRV, these descriptors would be defined by the light reflected from an object 
under a standard illuminant.
A further advance would be to develop methods for learning receptive fields that are
optimal for multivariate estimation problems.
With such methods one could directly estimate three-dimensional color descriptors.

@@

An important motivation for studying the computational problem of luminance constancy
is to gain insight about human vision.
The approach we developed can be used to make predictions of how humans would
perform in a psychophysical task that probes the visual representation of object LRV.
For example, one could study discrimination thresholds for target LRV using the
stimuli that were generated with the methods used
here.
We could ask how LRV discrimination thresholds are impacted by 
spectral variation in the illuminant, the background objects, and in the target object itself.
Comparing human thresholds with the precision of AMA-based estimates would then
allow inferences about how well human observers make use of the information available
in the images for performing the LRV discrimination task.
We think experiments and analyses along these lines represent an important future direction.
By coupling the experiments and analyses, we hope to advance our understanding
of how we see objects in the real world in a principled and controlled manner.

>>>>

Figure~\ref{fig:barGraphs} summarizes our main findings.
It plots the overall relative root mean squared estimation error (relative RMSE) for each condition. 
In Condition 1, where only the relative target object spectrum varies, luminance constancy poses an easy computational problem,
and performance is essentially perfect for both the baseline linear regression and AMA methods.
In Condition 2, where the spectral power distribution of the illumination also varies, the problem of luminance constancy is more difficult.
Indeed, neither of our methods performs well when they directly act on the cone excitations.
Performance with both methods improves greatly when the cone excitations are contrast normalized. 
Indeed, AMA performance for this case is, as in Condition 1, essentially perfect.
Linear regression does less well, in part because it only has access to cone excitations at image locations
corresponding to the target object.
Condition 2, however, does not include scene-to-scene variation in the background surface reflectances,
as occurs in natural viewing.
The results for Condition 3 show performance when we add such variation.
Performance here represents the overall level of luminance constancy that our methods achieve for the
most realistic condition that we tested.
In this case, AMA-based estimates of target object LRV are, on average, within 13\% (relative RMSE) of the true value.
Note that introduction of variation in the background surfaces makes the luminance constancy problem substantially harder,
because variation in the background surfaces reduces the reliability of the light reflected from these
surfaces as a cue to the spectral power distribution of the illuminant.

@@

As we noted in the introduction, supervised learning requires large labeled datasets.
These are not readily available  for the study of how to estimate correlates of object surface reflectance.
In particular, to obtain ground truth information about the reflectance of objects in a natural image, it is necessary to
make an independent of the illumination impinging on each object.
Although this can be done by inserting a small number of discretely spaced reflectance standards into the scene, 
interpolating to locations between such standards requires strong assumptions.
The use of rendered images allows us to work with large numbers of labeled images where object reflectance
is precisely known at each pixel.

A second advantage of using rendered images is that we can control the variation in distinct
scene factors that might affect the difficulty of the estimation problem.
This allows study of the individual effect of each factor as well as their combined effect.
Here, we exploited this to systematically introduce variation in the relative target object reflectance
spectrum, the spectrum of the illumination, and the reflectance spectrum of the background objects.
This allowed us to quantify how each of these variations limits LRV estimation.
We believe the basic approach can be extended to include parametric control over the amount
of variation of different factors as well as to include variation in additional factors (see below).
%% DHB: Update if we add additional analyses as per some of the comments above.

At the same time, it is important to recognize limitations of the rendered image approach.
Virtual rendered scenes are not guaranteed to capture all of the task relevant
variation that occurs in real scenes.
Indeed, even casual inspection of our images reveals that they are rendered and not real.
Computer graphics is getting better, however, and we expect that the gap between
virtual and natural image databases will steadily close.
Indeed, carefully constructed graphics images are now quite difficult to differentiate
from real images.
%% DHB: Might be able to reference Farid's recent book on this point.

Our use of virtual images relies on datasets that characterize the statistical variation
in object reflectance and natural illumination spectra.
We believe these datasets provide reasonable approximations.
None-the-less, they can be extended and improved.
For example, the dataset of illuminations does not include the artificial illuminations
that we now live much of our lives under.
The surface reflectance datasets capture constraints on naturally occurring reflectance
functions, but the surfaces in the datasets were not sample the environment in a manner that
characterizes the relative frequency of different reflectance spectra in natural viewing.
Future work with virtual image datasets will benefit from targeted efforts to characterize
the natural statistics of the scene components that are varied.
Under the assumption that combination of different scene factors (e.g. surface reflectances
and illuminations) is independent, the virtual image approach modularizes the study
of variation in natural scenes.
%Supervised learning requires large labeled data sets.  Such data sets are not readily available for the study of color constancy. Although there are databases of calibrated color images, these do not provide ground truth information about surface reflectance and illuminantion at each image location \cite{ChakrabartiHyperspectral,NascimentoFoster2016,ParragaHyperspectralData,TkacikUpennHypersepctralData,skauli2013collection,olmos2004biologically}. There are some smaller databases consisting of images of posed scenes where object surface reflectances are measured individually (\citeNP{funt1988color,ciurea2003large}; \citeauthor{davidPennHyperspectral}), but these are not large enough to drive supervised learning.

\subsection{Future Directions}
In the work presented here, we incorporated spectral variation into our statistical model
of naturalistic scenes.
It is natural to begin with spectral variation, because this variation is at the heart of
what makes luminance constancy a rich computational problem.
In natural scenes, however, there are other sources of variation that add additional
richness.
These include variation in non-spectral properties of objects and lighting in the scene, including
object texture, material, and shape as well as lighting geometry.
The methods we developed here may be generalized to study the effects of variation
in these factors.
The basic approach would be to incorporate the variation into the generation of the
scenes and again learn estimators from the corresponding labeled images.
A challenge for this approach will be to thoughtfully control the increase in problem
complexity, both to keep compute time feasible and to ensure that it is possible to
extract meaningful insight from the results.

Luminance constancy is a special case of the more general problem of color constancy.
The approach we developed here could be generalized to the estimation of additional
surface reflectance descriptors, such as object hue or chroma.
As with LRV, these descriptors would be defined by the light reflected from an object's
surface   under a standard illuminant.
A further advance would be to develop methods for learning receptive fields that are
optimal for multivariate estimation problems.
With such methods one could directly estimate three-dimensional color descriptors.

@@

An important motivation for studying the computational problem of luminance constancy
is to gain insight about human vision.
The approach we developed can be used to make predictions of how humans would
perform in a psychophysical task that probes the visual representation of object LRV.
For example, one could study discrimination thresholds for target LRV using the
stimuli that were generated with the methods used here.
More specfically, we could ask how LRV discrimination thresholds are impacted by 
spectral variation in the illuminant, the background objects, and in the target object itself.
Comparing human thresholds with the precision of AMA-based estimates would then
allow inferences about how well human observers make use of the information available
in the images for performing LRV discrimination.
We think experiments and analyses along these lines represent an important future direction.
By coupling the experiments and analyses we hope to advance, in a principled and controlled manner,
our understanding of how we see objects in the real world .
